Global patterns of forest autotrophic carbon fluxes

### 2020-06-05 manuscript/manuscript draft.Rmd
edited abstract
removed FACF (#80)
other minor edits

\newpage
###Abstract
Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
However, we lack a cohesive synthesis on how forest autotrophic carbon fluxes (FACF) vary globally with respect to climate and one another. 
Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant FACF, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
FACF generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).  **BBL: not sure the second part of this sentence merits inclusion in the abstract**
Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of FACF on this global scale. 
The effects of temperature were modified by moisture availability, with FACF reduced under hot and dry conditions and sometimes under very high precipitation. 
FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
Within the growing season, the influence of climate on C cycling was small but significant for a number of FACF. 
These findings clarify how FACF varies with latitude and climate on a global scale. 
As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

>>>>

\newpage
###Abstract
Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all flux variables analyzed, C cycling decreases linearly with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
C flux variables generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale.
**Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed**, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of C flux on this global scale. 
The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
These findings clarify how forest C flux varies with latitude and climate on a global scale. 
As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
